---
title: "Nurse Survey EDA"
format: 
  html:
    toc: true
    toc-depth: 2
    toc-title: TOC
    toc-location: left
  
editor: visual
execute: 
  message: false
  warning: false
---

## Packages

```{r}
#| message: false
#| warning: false

 
library(tidyverse)

library(corrplot)
```

Some Sources:

<https://r4ds.had.co.nz/exploratory-data-analysis.html#cat-cont>

<https://learningstatisticswithr.com/book/chisquare.html>

## Data Import

```{r}
data <- read_csv('./data/surveydata.csv')

data <- data %>% mutate(across(-c(ResponseID, Age), factor))

data
```


# Variables to investigate

We want to investigate the following by creating graphs and running chi squared on:

- Condom Use vs. App Sex Partners
- Condom Use vs. Total Sex Partners
- UT Health Services vs. Frequency of Testing
- UT Health Services vs. App Sex Partners
- UT Heatlth Services vs. Total Sex Partners
- Frequency of Testing vs. App Sex Partners
- Frequency of Testing vs. Total Sex Partners


## UT Services vs Sex Partners vs. App Sex Partners

```{r}
data %>%
  count(UTHC_serv, SexPart_APP, TOT_Sexpart)
```

```{r}
data %>%
  ggplot()+
  geom_bar(aes(x= UTHC_serv), color = 'white')+
  facet_grid(SexPart_APP ~ TOT_Sexpart)+
  theme_bw()
  
```

## UT Health Services vs. Sexual Partners on App

```{r}
data %>% 
  mutate(sexpartapp = case_when(SexPart_APP == '0' ~ '0',
                                SexPart_APP == '1' ~ '1',
                                SexPart_APP == '7' | SexPart_APP == '8' ~ '2+')
         ) %>%
  na.omit() %>% 
  count(UTHC_serv, sexpartapp)


```

```{r}
hcapp <- data %>% 
  count(UTHC_serv, SexPart_APP) %>% 
  na.omit() %>% 
  mutate(sexpartapp = case_when(SexPart_APP == '0' ~ '0',
                                SexPart_APP == '1' ~ '1',
                                SexPart_APP == '7' | SexPart_APP == '8' ~ '2+')
         ) %>%
  select(-SexPart_APP) %>%
  ggplot(aes(x= UTHC_serv, y = n, fill = sexpartapp))+
  geom_bar(stat = 'identity', color = 'white', position = 'dodge')+
  geom_text(aes(label = n), vjust = -1, colour = "black", position = position_dodge(.9))+
  #facet_grid(SexPart_APP ~ TOT_Sexpart)+
  #geom_point(aes(x = 2014, y = 440000), color = 'gold', fill = 'gold', shape = 23, size = 5)+
  #geom_text(aes(x = 2000, y = 447000, label = '2505 Bird'))+
  labs(title = 'Usage of UTA Health Services by Number of Sexual Partners Met Through Apps',
       x = '',
       y = 'Count',
       fill = '# Sexual Partners \n Through Apps')+
  scale_y_continuous(limits = c(0,70), breaks = seq(0,70, by = 5))+
  scale_x_discrete(labels = c('0' = 'Does Not Use UTA Health Service', '1' = 'Uses UTA Health Service') )+
    theme_minimal()+
  theme(plot.title = element_text(size = 12, face = 'bold'), 
        legend.position = 'right', 
        legend.text = element_text(size = 12), 
        #legend.title = element_blank(), 
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        axis.text.x = element_text(angle = 5), 
        #axis.text.y = element_text(angle = 15), 
        #axis.title.y = element_blank(), 
        #strip.text.x = element_text(size = 8, face = 'bold'),
        #legend.key.width = unit(0.1, 'in'), 
        #legend.key.height = unit(0.1,'in'), 
        #legend.margin = margin(t = 0, unit = 'cm'), 
        #plot.margin = unit(x = c(0.5,0.5,0.5,0.5), units = 'mm'), 
        #axis.ticks.y = element_blank()
        )

hcapp
```

## UT Health Services vs. Total Partners on App

```{r}
data %>% 
    mutate(TOT_Sexpart = as.numeric(TOT_Sexpart)) %>%
    mutate(sexparttot = case_when(TOT_Sexpart == '0' ~ '0',
                                TOT_Sexpart == '1' ~ '1',
                                TOT_Sexpart >= 1 ~ '2+') 
    ) %>%
  count(UTHC_serv, sexparttot) %>% 
  na.omit() 

         
```

```{r}
hctot <- data %>% 
    mutate(TOT_Sexpart = as.numeric(TOT_Sexpart)) %>%
    mutate(sexparttot = case_when(TOT_Sexpart == '0' ~ '0',
                                TOT_Sexpart == '1' ~ '1',
                                TOT_Sexpart >= 1 ~ '2+') 
    ) %>%
  count(UTHC_serv, sexparttot) %>% 
  na.omit()  %>%
  ggplot(aes(x= UTHC_serv, y = n, fill = sexparttot))+
  geom_bar(stat = 'identity', color = 'white', position = 'dodge')+
  geom_text(aes(label = n), vjust = -1, colour = "black", position = position_dodge(.9))+
  #facet_wrap(~ TOT_Sexpart)+
  #geom_point(aes(x = 2014, y = 440000), color = 'gold', fill = 'gold', shape = 23, size = 5)+
  #geom_text(aes(x = 2000, y = 447000, label = '2505 Bird'))+
  labs(title = 'Usage of UTA Health Services by Total Number of Sexual Partners',
       x = '',
       y = 'Count',
       fill = 'Total # of \n Sexual Partners')+
  scale_y_continuous(limits = c(0,70), breaks = seq(0,70, by = 5))+
  scale_x_discrete(labels = c('0' = 'Does Not Use UTA Health Service', '1' = 'Uses UTA Health Service') )+
    theme_minimal()+
  theme(plot.title = element_text(size = 12, face = 'bold'), 
        legend.position = 'right', 
        legend.text = element_text(size = 12), 
        #legend.title = element_blank(), 
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        axis.text.x = element_text(angle = 5), 
        #axis.text.y = element_text(angle = 15), 
        #axis.title.y = element_blank(), 
        #strip.text.x = element_text(size = 8, face = 'bold'),
        #legend.key.width = unit(0.1, 'in'), 
        #legend.key.height = unit(0.1,'in'), 
        #legend.margin = margin(t = 0, unit = 'cm'), 
        #plot.margin = unit(x = c(0.5,0.5,0.5,0.5), units = 'mm'), 
        #axis.ticks.y = element_blank()
        )

hctot
```



## Condom Use vs. Total Sexual Partners

```{r}
data %>% 
    mutate(TOT_Sexpart = as.numeric(TOT_Sexpart)) %>%
    mutate(sexparttot = case_when(TOT_Sexpart == '0' ~ '0',
                                TOT_Sexpart == '1' ~ '1',
                                TOT_Sexpart >= 1 ~ '2+') 
    ) %>%
  count(CU_lastSex, sexparttot) %>% 
  na.omit() 

         
```

```{r}
cutot <- data %>% 
    mutate(TOT_Sexpart = as.numeric(TOT_Sexpart)) %>%
    mutate(sexparttot = case_when(TOT_Sexpart == '0' ~ '0',
                                TOT_Sexpart == '1' ~ '1',
                                TOT_Sexpart >= 1 ~ '2+') 
    ) %>%
  count(CU_lastSex, sexparttot) %>% 
  na.omit()  %>%
  ggplot(aes(x= CU_lastSex, y = n, fill = sexparttot))+
  geom_bar(stat = 'identity', color = 'white', position = 'dodge')+
  geom_text(aes(label = n), vjust = -1, colour = "black", position = position_dodge(.9))+
  #facet_wrap(~ TOT_Sexpart)+
  #geom_point(aes(x = 2014, y = 440000), color = 'gold', fill = 'gold', shape = 23, size = 5)+
  #geom_text(aes(x = 2000, y = 447000, label = '2505 Bird'))+
  labs(title = 'Condom Usage by Total Number of Sexual Partners',
       x = '',
       y = 'Count',
       fill = 'Total # of \n Sexual Partners')+
  scale_y_continuous(limits = c(0,70), breaks = seq(0,70, by = 5))+
  scale_x_discrete(labels = c('0' = 'Did Not Use Condom \n Last Sexual Encounter', '1' = 'Used Condom \n Last Sexual Encounter') )+
    theme_minimal()+
  theme(plot.title = element_text(size = 12, face = 'bold'), 
        legend.position = 'right', 
        legend.text = element_text(size = 12), 
        #legend.title = element_blank(), 
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        axis.text.x = element_text(angle = 5), 
        #axis.text.y = element_text(angle = 15), 
        #axis.title.y = element_blank(), 
        #strip.text.x = element_text(size = 8, face = 'bold'),
        #legend.key.width = unit(0.1, 'in'), 
        #legend.key.height = unit(0.1,'in'), 
        #legend.margin = margin(t = 0, unit = 'cm'), 
        #plot.margin = unit(x = c(0.5,0.5,0.5,0.5), units = 'mm'), 
        #axis.ticks.y = element_blank()
        )

cutot
```

## Condom Use vs. App Sexual Parnters

```{r}
data %>% 
    mutate(sexpartapp = case_when(SexPart_APP == '0' ~ '0',
                                SexPart_APP == '1' ~ '1',
                                SexPart_APP == '7' | SexPart_APP == '8' ~ '2+')
    ) %>%
  count(CU_lastSex, sexpartapp) %>% 
  na.omit()
```

```{r}
cuapp <- data %>% 
    mutate(sexpartapp = case_when(SexPart_APP == '0' ~ '0',
                                SexPart_APP == '1' ~ '1',
                                SexPart_APP == '7' | SexPart_APP == '8' ~ '2+')
    ) %>%
  count(CU_lastSex, sexpartapp) %>% 
  na.omit() %>%
  add_row(CU_lastSex = '0', sexpartapp = '2+', n = 0) %>%
  ggplot(aes(x= CU_lastSex, y = n, fill = sexpartapp))+
  geom_bar(stat = 'identity', color = 'white', position = 'dodge')+
  geom_text(aes(label = n), vjust = -1, colour = "black", position = position_dodge(.9))+
  #facet_wrap(~ TOT_Sexpart)+
  #geom_point(aes(x = 2014, y = 440000), color = 'gold', fill = 'gold', shape = 23, size = 5)+
  #geom_text(aes(x = 2000, y = 447000, label = '2505 Bird'))+
  labs(title = 'Condom Usage by Number of Sexual Partners Met Through Apps',
       x = '',
       y = 'Count',
       fill = '# of Sexual Partners \n Met Through Apps')+
  scale_y_continuous(limits = c(0,70), breaks = seq(0,70, by = 5))+
  scale_x_discrete(labels = c('0' = 'Did Not Use Condom \n Last Sexual Encounter', '1' = 'Used Condom \n Last Sexual Encounter') )+
    theme_minimal()+
  theme(plot.title = element_text(size = 12, face = 'bold'), 
        legend.position = 'right', 
        legend.text = element_text(size = 12), 
        #legend.title = element_blank(), 
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        axis.text.x = element_text(angle = 5), 
        #axis.text.y = element_text(angle = 15), 
        #axis.title.y = element_blank(), 
        #strip.text.x = element_text(size = 8, face = 'bold'),
        #legend.key.width = unit(0.1, 'in'), 
        #legend.key.height = unit(0.1,'in'), 
        #legend.margin = margin(t = 0, unit = 'cm'), 
        #plot.margin = unit(x = c(0.5,0.5,0.5,0.5), units = 'mm'), 
        #axis.ticks.y = element_blank()
        )

cuapp
```

# Graphs

```{r}
hcapp
hctot
cuapp
cutot

ggsave(plot = hcapp, filename = './results/figures/healthserveapp.png', dpi = 320)
ggsave(plot = hctot, filename = './results/figures/healthservetotal.png', dpi = 320)
ggsave(plot = cuapp, filename = './results/figures/condomuseapp.png', dpi = 320)
ggsave(plot = cutot, filename = './results/figures/condomusetotal.png', dpi = 320)
```

# Chi Squared Tests

This is where we evaluate the chi squared independance test. Results should be written in the paper using this APA format (https://www.socscistatistics.com/tutorials/chisquare/default.aspx)

The chisq test in R returns:

**Statistic:** the value the chi-squared test statistic. parameter: the degrees of freedom p.value: the p-value of the test observed: the observed count expected: the expected count

P value less than 0.05 and a chi-squared value higher than the critical value reject null hypothesis that there is no relationship.

Sources:

<http://www.sthda.com/english/wiki/chi-square-goodness-of-fit-test-in-r> <http://www.sthda.com/english/wiki/wiki.php?title=chi-square-test-of-independence-in-r>

## Chi Squared Tests: Condom Use vs. App Sexual Partners

```{r}

data_cuapp <- data %>% 
  count(CU_lastSex, SexPart_APP) %>% na.omit() %>%
  pivot_wider(names_from = SexPart_APP, values_from = n) %>% 
  replace(is.na(.),0) %>% 
  column_to_rownames(var = 'CU_lastSex')

data_cuapp 

data %>% 
  count(CU_lastSex, SexPart_APP) 

```

```{r}
chisq_cuapp <- chisq.test(data_cuapp, simulate.p.value = TRUE)

chisq_cuapp

chisq_cuapp$observed

chisq_cuapp$expected %>% round(2)

print('Crtical Value of 95% Chi Square with 3 DOF:')
qchisq(p = 0.975, df = 3)

```


```{r}
corrplot(chisq_cuapp$residuals, is.cor = FALSE)


chisq_cuapp$residuals
```

Positive residuals are in blue.

Positive values in cells specify an attraction (positive association) between the corresponding row and column variables.

Negative residuals are in red.

This implies a repulsion (negative association) between the corresponding row and column variables.

## Chi Squared Tests: Condom Use vs. Total Sexual Partners

```{r}

data_cutot <- data %>% 
  count(CU_lastSex, TOT_Sexpart) %>% na.omit() %>%
  pivot_wider(names_from = TOT_Sexpart, values_from = n) %>% 
  replace(is.na(.),0) %>% 
  column_to_rownames(var = 'CU_lastSex')

data_cutot 

```

```{r}
chisq_cutot <- chisq.test(data_cutot)

chisq_cutot

chisq_cutot$observed

chisq_cutot$expected %>% round(2)

print('Crtical Value of 95% Chi Square with 6 DOF:')
qchisq(p = 0.975, df = 6)

```

The chisq test returns:

**Statistic:** the value the chi-squared test statistic. parameter: the degrees of freedom p.value: the p-value of the test observed: the observed count expected: the expected count

P value less than 0.05 and a chi-squared value higher than the critical value reject null hypothesis that there is no relationship.

```{r}
corrplot(chisq_cutot$residuals, is.cor = FALSE)


chisq_cutot$residuals
```

## Z Proportion and Chi Squared Tests: UT Health Services vs. App Partners

Note that, for 2 x 2 table, the standard chi-square test in chisq.test() is exactly equivalent to prop.test() but it works with data in matrix form.

Source:

<http://www.sthda.com/english/wiki/two-proportions-z-test-in-r>

```{r}
data_cuapp

data_cuapp1 <- data %>% 
   mutate(sexpartapp = case_when(SexPart_APP == '0' ~ '0',
                                SexPart_APP != '0' ~ '1+')
    ) %>%
  count(CU_lastSex, sexpartapp) %>% 
  na.omit() %>%
  pivot_wider(names_from = sexpartapp, values_from = n) %>% 
  replace(is.na(.),0) %>% 
  column_to_rownames(var = 'CU_lastSex')


data_cuapp1
```


```{r}
chisq_cuapp1 <- chisq.test(data_cuapp1)

chisq_cuapp1

chisq_cuapp1$expected

prop.test(as.matrix(data_cuapp1), alternative = "two.sided",
          correct = TRUE)

```


## Z Proportion and Chi Squared Tests: UT Health Services vs. Total Partners

```{r}
data_hctot <- data %>% 
    mutate(sexparttot = case_when(TOT_Sexpart == '1' ~ '1',
                                TOT_Sexpart != '1' ~ '2+')    ) %>%
  count(UTHC_serv, sexparttot) %>% 
  na.omit() %>%
  pivot_wider(names_from = sexparttot, values_from = n) %>%
  column_to_rownames(var = 'UTHC_serv')

data_hctot
```

Shows that the proportion test for a 2 x 2 table is the same as a Pearson Chi Squared Test

```{r}
prop.test(as.matrix(data_hctot), alternative = "two.sided",
          correct = TRUE)

chisq.test(data_hctot)
```


# Misc Explorations

## Type of Sex vs. Condom Use vs. Substance Use

```{r}
data %>% count(SubUse_lastSex, CU_lastSex)
```

```{r}
data %>% 
  ggplot()+
  geom_bar(aes(x= Or_VagAnal), color = 'white')+
  facet_grid(CU_lastSex ~ SubUse_lastSex)+
  theme_bw()
```

## STI vs. Condom

```{r}
data %>% count(CU_lastSex, Pos_STIHIV)
```

```{r}
data %>%
  ggplot()+
  geom_bar(aes(x = Pos_STIHIV))+
  facet_wrap(~ CU_lastSex)
```

## Prefer Plan B vs. Condom

```{r}

data %>% count(Prefer_planB, CU_lastSex)
```

```{r}
data %>% 
  # count(Prefer_planB, CU_lastSex) %>%
  ggplot()+
  geom_bar(aes(x= Prefer_planB), color = 'white')+
  facet_wrap(~ CU_lastSex)+
  theme_bw()
```

```{r}

data %>% count(TOT_Sexpart, SexPart_APP)
```

```{r}
data %>% 
  # count(Prefer_planB, CU_lastSex) %>%
  ggplot()+
  geom_bar(aes(x= SexPart_APP), color = 'white')+
  facet_wrap(~ TOT_Sexpart)+
  theme_bw()
```



## Type of Sex vs. Race, Age, and More EDA Plots

```{r}
data %>% 
  count(Or_VagAnal, Race)
```

```{r}
data %>% 
  ggplot(aes(x = Age, y = ..density..))+
  geom_freqpoly(aes( color = Or_VagAnal), binwidth = 5)
```

```{r}
data %>% 
  ggplot(aes(x = Or_VagAnal, y = Age))+
  geom_boxplot()
```

```{r}
data %>% 
  ggplot()+
  geom_count(aes(x = Or_VagAnal, y = CU_lastSex))
```

```{r}
data %>% 
  count(Or_VagAnal, CU_lastSex) %>%
  ggplot()+
  geom_tile(aes(x = Or_VagAnal, y = CU_lastSex, fill = n))
```

```{r}
#| echo: false

data %>%
  ggplot(aes(x = Age, y = Or_VagAnal
             #, group = week, shape = week, color = week
             ))+
  geom_point()+
  geom_line()+
  #geom_point(aes(x = 2014, y = 440000), color = 'gold', fill = 'gold', shape = 23, size = 5)+
  #geom_text(aes(x = 2000, y = 447000, label = '2505 Bird'))+
  labs(title = 'EDA Plot',
       x = 'Age',
       y = 'Test Stat')+
  #scale_y_continuous(limits = c(0,110), breaks = seq(0,110, by = 10))+
  #scale_x_continuous(limits = c(1250,2500), breaks = seq(1250, 2500, by = 250))+
    theme_minimal()+
  theme(plot.title = element_text(size = 12, face = 'bold'), 
        legend.position = 'right', 
        legend.text = element_text(size = 12), 
        #legend.title = element_blank(), 
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        axis.text.x = element_text(angle = 15), 
        #axis.text.y = element_text(angle = 15), 
        #axis.title.y = element_blank(), 
        #strip.text.x = element_text(size = 8, face = 'bold'),
        #legend.key.width = unit(0.1, 'in'), 
        #legend.key.height = unit(0.1,'in'), 
        #legend.margin = margin(t = 0, unit = 'cm'), 
        #plot.margin = unit(x = c(0.5,0.5,0.5,0.5), units = 'mm'), 
        #axis.ticks.y = element_blank()
        )

```

```{r}
#| eval: false
#| echo: false


wc3 <- ggplot() + geom_bar(data = water_capacitydf_water, aes(x = Techtype, y = Capacity, fill = Technology), stat = 'identity')+
  facet_wrap(~Scenario)+
  #theme_minimal()+
  theme(plot.title = element_text(size = 12, face = 'bold'), legend.position = 'bottom', legend.text = element_text(size = 8), legend.title = element_blank(), axis.title = element_text(size = 8), axis.text = element_text(size = 8), axis.text.x = element_text(angle = 15), axis.text.y = element_blank(), axis.title.y = element_blank(), strip.text.x = element_text(size = 8, face = 'bold'),legend.key.width = unit(0.1, 'in'), legend.key.height = unit(0.1,'in'), legend.margin = margin(t = 0, unit = 'cm'), plot.margin = unit(x = c(0.5,0.5,0.5,0.5), units = 'mm'), axis.ticks.y = element_blank())+
  labs(title = 'Water Only', y = 'Average Installed Water Capacity (kGals)', x = '')+
   scale_fill_manual(labels = c("C_VRF" = "WRF", "H_RW" = "RWH", 'H_GW' = 'GWR', "U_H2O1" = 'U_H2O1', "U_H2O2" = 'U_H2O2', "U_H2O3" = 'U_H2O3', "U_H2O4" = 'U_H2O4', "U_H2O5" = 'U_H2O5','C_RWTANK' = 'CRWTANK', 'H_RWTANK' = 'RWTANK'), values = c("C_VRF" = "#E31A1C", "H_RW" = "#FB9A99", 'H_GW' = 'darkorange2', "U_H2O1" = 'gray45', "U_H2O2" = 'gray38', "U_H2O3" = 'gray32', "U_H2O4" = 'gray25', "U_H2O5" = 'gray15','C_RWTANK' = 'aquamarine', 'H_RWTANK' = 'aquamarine4') )+
  coord_cartesian(ylim = c(0,1.25))+
  guides(fill=guide_legend(nrow = 2, byrow = TRUE))
```
